EGGs are adhesive!

### 2025-03-13 calco/main.tex

\commentato{ 
\begin{restatable}{corollary}{rege}\label{cor:rege}
		An arrow $h\colon \mathcal{G \to H}$  is a regular mono in $\egg$ if and only if $I(h)$ is a regular mono in $\hyp$.  In particular a morphism of $\egg$ is a regular mono if and only if all its components are injections.
\end{restatable}
\begin{proof}
	contenuto...
\end{proof}
}

@@

>>>>

@@

\commentato{ 
\begin{remark}\label[remark]{rem:sig_egg}
	Given a signature $\Sigma = (O_\Sigma, \ari_\Sigma)$,
	the hypergraph 
	$R(\mathcal{G}^\Sigma) = (O_\Sigma, 1, 1, \ari_\Sigma, \delta_1, \id{1})$
	is an object of $\egg$.
	Indeed, if $(S, \pi_1, \pi_2)$ is the kernel pair of $\id{1}^\star \circ \ari_\Sigma$,
	we have that $\id{1}^\star \circ \delta_1 \circ \pi_1$ and $\id{1}^\star \circ \delta_1 \circ \pi_2$ are both the contant functions of value $1$.
\end{remark}

\begin{definition}\label[definition]{def:label_egraphs}
	Let $\Sigma$ be an algebraic signature, and $\mathcal{G}^\Sigma$ the hypergraph associated to it.
	A \emph{labeled e-hypergraph} is a pait $(\mathcal{E}, l)$, where $\mathcal{E}$ is an e-hypergraph and $l: T \circ I (\mathcal{H}) \to \mathcal{G}^\Sigma$ is a morphism of $\hyp$.
	A \emph{morphism of labeled e-hypergraphs} $(\mathcal{E}, l)\to (\mathcal{E}', l')$ is a morphism $h: \mathcal{E \to E}'$ such that $l = l' \circ (T \circ I)(h)$.
	We denote the resulting category by $\eggs$.
\end{definition}

%Functor from GEqhyp_S to EqHyp_S

With a similar argument as the one done in \Cref{prop:slice}, and by \Cref{rem:sig_egg} we get a characterization of $\eggs$.

\begin{proposition}
	The category $\eggs$ is equivalent to $\egg/R(\mathcal{G}^\Sigma)$
\end{proposition}

Consider now the functor that forgets the label $V_\eq : \eggs \to \egg$, and $\pbes$ the class of morphisms in $\eggs$ whose image through $V_\eq$ is in $\pbe$. From \Cref{thm:slice-functors}, \Cref{lim},  \Cref{lem:elim,lem:epo}, we can deduce what follows.

\begin{proposition}\label[proposition]{prop:eadh}
	The following facts hold true:
	\begin{enumerate}
	\item $\eggs$ has all limits and pushouts along regular monos, which are created by $V_\eq$;
	\item $\eggs$ is $\pbes$-adhesive.
	\end{enumerate}
\end{proposition}

Now, by \Cref{cor:rege,cor:mono1,cor:mono2}, we get what follows.

\begin{corollary}\label{cor:emon}
	Let $h = (h_E, h_V, h_Q)$ be an arrow in $\eggs$. Then the following holds.
	\begin{enumerate}
	\item $h$ is mono if and only if $h_E$ and $h_V$ are injective;
	\item $h$ is regular mono in and only if each component of it is injective.
	\end{enumerate}
\end{corollary}

We can now introduce term graphs with closed equivalences.

%\begin{definition}\label[definintion]{def:GEqTGs}
%	Given a signature $\Sigma$, we define the category $\GEqTGs$ as the full subcategory of $\eggs$ in which objects are arrows $l: \mathcal{G} \to R(\mathcal{G}^\Sigma)$ where $t_\mathcal{G}$ is mono. The inclusion functor is denoted $I_{\GEqTGs} : \GEqTGs \to \egg/R(\mathcal{G}^\Sigma)$.
%\end{definition}

\begin{definition}\label[definition]{def:GEqTGs}
	Let $\Sigma$ be an algebraic signature. An object $(\mathcal{T}, l)$ of $\eggs$ is an \emph{e-term graph} if $l: T(\mathcal{T})\to \mathcal{G}^\Sigma$ is a term graph. We denote by $\GEqTGs$ the full subcategory of $\eggs$ so defined, and denote by $I_{\GEqTGs}$ the corresponding inclusion functor.
\end{definition}
\noindent
\parbox{8.5cm}{\begin{remark}\label[remark]{rem:inclusions}
	The category $\GEqTGs$ is also a full subcategory of $\EqTGs$, since each object $(\mathcal{H}, l)$ is a term graph with equivalence too.
	We denote the corresponding inclusion functor as $S_\eq$. Hence, we have a situation similiar to the one in \Cref{rem:obv}, as the diagram aside shows, where $T_\eq$ is the inclusion functor of \Cref{def:ehyp1} lifted to the slice category, i.e., is such that $V_\Sigma \circ T_\eq = I \circ V_\eq$.
\end{remark}}
	\hfill\parbox{4.5cm}{\[\xymatrix@R=20pt@C=38pt{\GEqTGs \ar[r]^{I_{\GEqTGs}}  \ar[d]_{S_\eq}& \eggs \ar[d]^{T_\eq} \\ \EqTGs \ar[r]_{I_\eq}& \EqHyp_{\Sigma}}\]}

%\todo{Here we must introduce $\GEqTGs$}
>>>>>>> refs/remotes/origin/main}
